Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-04-29 experiments.tex
update exp

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
In this subsection we evaluating the behavior of the proposed approach
using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as a testbed. The same philosophy
behind a VLIW system makes it a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: these architectures let designers
easily trade between power, energy and performances~\cite{}, leading
to Pareto sets whose extension is an ideal ground for testing the
proposed Parameters space representation.
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
constitutes the key feature to enable the multi-objective trade-offs
of the point above. It is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario more suitable for investigating hardly predictable
hardware/software interactions, especially when compared to
architectures where, in order to preserve binary compatibility,
software generation processes are independent from the hardware that
will execute it.

System parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiling options}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type of opera. In particular, five different types of register files can be identified:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

Further, a set of the most impacting compilation parameters has been
selected and included in the configuration space:
\begin{itemize}
\item{tcc\_region}: Specifies the scope of action of the compiler and the type
of code transformation involved (basic block, super block and hyper
block) 
\item {max\_unroll\_allowed}: The number of unroll iterations allowed
\item{regroup\_only}: Avoids inlining 
\item{do\_classic\_opti}: A set of classical optimization, not VLIW related,
such as common expression removal 
\item{do\_prepass\_scalar\_scheduling}: Performs a schedule before
forming regions 
\item{do\_postpass\_scalar\_scheduling}: Performs a schedule after the region formation 
do\_modulo\_scheduling & Modular scheduling 
\item{memvr\_profiled}: Performs a memory-dependencies profiling 
\end{itemize}

@@

The results presented in the next section compare PS against a widely
adpted Multi-objective Genetic Algorithm~\cite{}. At best of our
knowledge, genetic approaches
still represent the best compromise between efficiency (i.e. time
required for exploration) and accuracy of the reported Paret Sets. Of
course mono-objective approaches have been
deliberately excluded from this analysis. Other
approaches~\cite{dep} that make use of some a-priori knowledge about the
role and meaning of each parameter have been also excluded, since our
aim is evaluate the capacity of the strategy to focus on interesting
parameter regions without any external judgement or help based on
euristics.  Finally, it should be pointed
out that in this introductory work we explicitly focus on TODO
remainding to works like~\cite{} for a detailed analysis of other
similar multi-objective strategies.

The class of benchmark being considered is representative of some common
frequently running application kernels in an embedded multimedia
environment. Table~\ref{tab:bench} shows the set of applications
chosen along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Fib & Fibonacci series \\
	Mm & Floating point matrices multiplication \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}
%------------------------------------------------------------------------------

\section{Results}

Each data set is obtained evaluation a 1,000 randomly chosen point of
the configuration space. In particular, two different scenarios have
been considered: the first one, referred as \emph{Variable Compilation
Profile (VCP)}, includes all compilation parameters described
in the Table~\ref{tab:compiler}; a second scenario, \emph{Fixed
Compilation Profile (FCP)}, which excludes any of those parameters
exploring all the remaining hardware aspects of the architecture as
described in Section~\ref{subsec:arch}.

@@

Let us now analyse the Pareto fronts from a quantitative viewpoint. We
consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:variation_range}. As it can
be observed, the VCP exploration provides solutions which fall on a
range that is, on average, 23\% and 40\% wider than that provided by a
FCP exploration for power dissipation and execution time,
respectively.
\begin{figure}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps}
  \end{tabular}
  \caption{Variation range for different benchmarks between a FCP and
    a VCP exploration for power dissipation and execution time.}
  \label{fig:variation_range}
\end{figure}

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances~\cite{}, leading
to Pareto sets whose extension is an ideal ground for our testing
purposes focused on parameter-space space representation of Paretos.
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
constitutes the key feature to enable the multi-objective trade offs
of the point above. It is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario well suitable for investigating hardly predictable
hardware/software interactions, especially when compared to
architectures that, in order to preserve binary compatibility,
separate software generation processes from the hardware that
will execute it.

System parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

Next, a set of the most impacting code compilation parameters has been
selected and included in the configuration space:
\begin{itemize}
\item{tcc\_region}: Specifies the scope of action of the compiler and the type
of code transformation involved (basic block, super block and hyper
block) 
\item {max\_unroll\_allowed}: The number of unroll iterations allowed
\item{regroup\_only}: Avoids inlining 
\item{do\_classic\_opti}: A set of classical optimization, not VLIW related,
such as common expression removal 
\item{do\_prepass\_scalar\_scheduling}: Performs a schedule before
forming regions 
\item{do\_postpass\_scalar\_scheduling}: Performs a schedule after the region formation 
do\_modulo\_scheduling & Modular scheduling 
\item{memvr\_profiled}: Performs a memory-dependencies profiling 
\end{itemize}

@@

The results presented in the next section compare PS against a widely
adpted Multi-objective Genetic Algorithm~\cite{}. At best of our
knowledge, genetic approaches
still represent the best compromise between efficiency (i.e. time
required for exploration) and accuracy of the reported Paret Sets. Of
course, mono-objective approaches have been
deliberately excluded from this analysis. Other
approaches~\cite{dep} that make use of some a-priori knowledge about the
role and meaning of each parameter have been also excluded, since our
aim is evaluate the capacity of focusing on interesting
parameter regions without any external judgement or help based on
euristics.  Finally, it should be pointed
out that in this introductory work we will explicitly use the TODO
variant of MOGA, remainding to works like~\cite{} for a detailed
analysis on how it differs and what could be expected from similar multi-objective strategies.

The class of benchmark being considered is representative of some common
frequently running application kernels in an embedded multimedia
environment. Table~\ref{tab:bench} shows the set of applications
chosen along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}
%------------------------------------------------------------------------------

\subsection{Results}

Set of Figures~\ref{fig:pareto_fronts_100}
and~\ref{fig:pareto_fronts} show the resultig pareto fronts for two
different scenarios of 100 and 500 simulations budget.

@@

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:variation_range}. As it can
be observed, the VCP exploration provides solutions which fall on a
range that is, on average, 23\% and 40\% wider than that provided by a
FCP exploration for power dissipation and execution time,
respectively.
\begin{figure}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps}
  \end{tabular}
  \caption{Variation range for different benchmarks between PS and GA
  exploration for power dissipation and execution time.}
  \label{fig:variation_range}
\end{figure}
